Generalized Tractability

### 2023-06-04 alternative.tex
Updates from Overleaf

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=L_{(p,q)} d^{q}}^\infty (\min\{1,\lambda_{i,d}^p\})  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\begin{example}[Strong exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where $\vp$ is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for strong exponential polynomial tractability \[
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty (\min\{1,\log(1+\lambda_{i,d}\})^{p}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals
\] In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=L_{(p,q)} d^{q}}^\infty [\min\{1,\log(1+\lambda_{i,d}\})]^p < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
\end{example}

\begin{example}[(Strong) Separable tractability]
Suppose that the tractability function is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$, hence $T$ is non-decreasing in all variables.

We say that a problem is defined as $T$-separable tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp}\,\phi(\varepsilon^{-1},p)\, \psi(d,q)\qquad \forall \varepsilon >0, \, d\in\naturals.
\]
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}\phi(\varepsilon^{-1},p)\psi(1,q) \qquad \forall \varepsilon >0, \, d\in\naturals,
\] where $\vp = (p,q)$. 
Furthermore, we assume the following as in \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.
\[
\lim_{\varepsilon\rightarrow0}\phi(\varepsilon^{-1},p) = \infty \qquad \forall d\in\naturals, p >0.
\]
Also since $T(\cdot,d,\vp)$ is non-decreasing, we define the following limit
\[
T(0,d,\vp):= \lim_{\varepsilon \rightarrow \infty}T(\varepsilon^{-1},d,\vp)\geq T(0,1,\textbf{0})>0
\] 
Additionally, we have the following technical assumptions. Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, we have that 
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau_1} &\le \hat{K}_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\
      [\psi(d,q)]^{\tau_2} &\le \hat{K}_{q,\tau_2} \psi (d,\tau_2 q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals.
\end{align*} where $L_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, and where $L_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.
\begin{align*}
  [T(\varepsilon^{-1},d,\vp]^\tau \leq K_{\vp,\tau}T(\varepsilon^{-1},d,\tau\vp) \quad \forall \varepsilon \in (0,\infty),\, d\in \mathbb{N},\, \vp\in[0,\infty)^2, \tau >0\end{align*}

Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = L_{(p,q)\psi(d,q)}}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\]
%As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated.

\end{example}

\begin{example}[Non-separable tractability]
    Let the tractability function $T$, be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(\phi(\varepsilon^{-1})\, \psi(d))\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$ and $\vp = p$. 
We say that a problem is non-separable tractable with parameter $\vp$ iff there exists a positive $C_{\vp}$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp} T(\varepsilon^{-1},d,\vp),
\]
for all $\varepsilon\in (0,1]$ and $d\in \naturals$, that is
\[\comp(\varepsilon,d) \leq  \exp\{p(\phi(\varepsilon^{-1})\psi(d))\}.\]

Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for non-separable tractability
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,\psi(d)} \phi(\lambda_{i,d})\psi(d) < \infty \qquad \text{for some} p \geq 0, L_p \in \naturals.
\]
We have the following special cases. \end{example}
\begin{example}[Algebraic quasi-polynomial tractability] Here we consider another common notions of tractability.

@@

\noindent Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\frac{1}{d^p}\sum_{i=Lp\,d^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)} < \infty \qquad \text{for some } p \geq 0, L_p \in \naturals.\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

@@

%In the following we introduce two specific cases of tractability which depend on the form of the function $T$ (see \cite{}).

>>>>

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_{(p,q)} d^{q}}^\infty (\min\{1,\lambda_{i,d}^p\}) d^{-q} < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\begin{example}[Strong exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where $\vp$ is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for strong exponential polynomial tractability \[
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty (\min\{1,\log(1+\lambda_{i,d}\})^{p}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals
\] In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability
\[
 \sup_{d\in\naturals} \sum_{i=L_{(p,q)} d^{q}}^\infty [\min\{1,\log(1+\lambda_{i,d}\})]^p d^{-q} < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
\end{example}

\begin{example}[Algebraic quasi-polynomial tractability] Here we consider another common notions of tractability.

@@

\noindent Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,e^pd^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)}e^{-p}d^{-p} < \infty \qquad \text{for some} p \geq 0, L_p \in \naturals.
\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

@@

In the following we introduce two specific cases of tractability which depend on the form of the function $T$ (see \cite{}).
\begin{example}[(Strong) Separable tractability]
Suppose that the tractability function is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$, hence $T$ is non-decreasing in all variables.

We say that a problem is defined as $T$-separable tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp}\,T(\varepsilon^{-1},d,\vp)\qquad \forall \varepsilon >0, \, d\in\naturals,
\]
that is
\[
\comp(\varepsilon,d) \leq \phi(\varepsilon^{-1},p)\, \psi(d,q) \qquad \forall \varepsilon >0, \, d\in\naturals.
\]
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}T(\varepsilon^{-1}, 1,\vp) \qquad \forall \varepsilon >0, \, d\in\naturals,
\] where $\vp = p$. That is, we have
\[
\comp(\varepsilon,d) \leq \phi(\varepsilon^{-1},p)\qquad \forall \varepsilon >0, \, d\in\naturals.
\]
Furthermore, we assume the following as in \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.
\[
\lim_{\varepsilon\rightarrow0}T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d\in\naturals, \vp \in[0,\infty)^2.
\]
Also since $T(\cdot,d,\vp)$ is non-decreasing, we define the following limit
\[
T(0,d,\vp):= \lim_{\varepsilon \rightarrow \infty}T(\varepsilon^{-1},d,\vp)\geq T(0,1,\textbf{0})>0
\] 
{\color{red} Additionally, we have the following technical assumptions. Suppose that, for any real $\tau>0$ we have
\begin{align*}
  [T(\varepsilon^{-1},d,\vp)]^\tau \leq K_{\vp,\tau}T(\varepsilon^{-1},d,\tau\vp) \quad \forall \varepsilon \in (0,\infty),\, d\in \mathbb{N},\, \vp\in[0,\infty)^2, \tau >0,\end{align*} that is for positive $K_{\vp,\tau}$ we have by definition
  \[
  [\phi(\varepsilon^{-1},p)\, \psi(d,q) ]^\tau \leq K_{\vp,\tau} \phi(\varepsilon^{-1},\tau p)\, \psi(d,\tau q) \]
%Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly separable tractability
%\begin{equation*} \label{eq:strong_tractiff}    \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = L_{\vp\psi(d,q)}}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\]
Observe that the only place where assumption (\ref{eq:ptauassume}) is used is in the necessary condition as follows.
Suppose that we have
$T$-separable tractability. That is, for some $\vp \ge \vzero$, there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\]
Since the sequence of eigenvalues $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K4}
	\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\end{equation}
Define the integers
\[
i (\varepsilon,d,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1, \qquad
i (\infty,d,\vp) := \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1 \ge 1.
\]
Thus, it follows by \eqref{eq:lambda_K4} that $\lambda_{i(\varepsilon,d,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,d,\vp)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,d,\vp),d}^{-1},d,\vp)+1, \qquad \forall \varepsilon > 0, \ d \in \naturals,
\]
since
$T(\cdot,d,\vp)$ is non-decreasing.

For $\varepsilon$ taking on all positive values, $i(\varepsilon,d,\vp)$ takes on all values greater than or equal to $i(\infty,d,\vp)$, so
\[
i\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i,d}^{-1},d,\vp)+1, \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals.
\] That is
\[
i \leq C_{\vp}\phi(\varepsilon^{-1},p)\psi(d,q)+1\leq C_{\vp}\phi(\lambda^{-1}_{i,d},p)\psi(d,q)+1
\]
This implies via our technical assumption \eqref{eq:ptauassume} that for any $\tau > 1$,
\begin{gather*}
	K_{\vp,\tau}\,\phi (\lambda_{i,d}^{-1},\tau p)\psi(d,\tau q) \ge
	[\phi(\lambda_{i,d}^{-1}, p)\psi(d,q)]^\tau
	\ge
	\left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals, \\
	 \frac{1}{\phi (\lambda_{i,d}^{-1},\tau p)\psi(d,\tau q) } \le
	\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,\vp),2\},  \ d \in \naturals.
\end{gather*}

Let $\vp' = (\tau p,\tau q)$.  Note that $i(\infty,d,\cdot)$ is non-decreasing.  Summing both sides of the latter inequality over $i$ from $i(\infty,d,\vp') + 1 \ge i(\infty,d, \vp) + 1 \ge 2$ to $\infty$ it follows that
\begin{equation*}
	\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')}
	 \le  K_{\vp,\tau}\, C_{\vp}^\tau
	\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
	 \le  K_{\vp,\tau}\, C_{\vp}^\tau
	\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.

%As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated. 
}
\end{example}
\begin{example}[Non-separable tractability]
    Let the tractability function $T$, be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(\phi(\varepsilon^{-1})\, \psi(d))\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$ and $\vp = p$. 
We say that a problem is non-separable tractable with parameter $\vp$ iff there exists a positive $C_{\vp}$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp} T(\varepsilon^{-1},d,\vp),
\]
for all $\varepsilon\in (0,1]$ and $d\in \naturals$, that is
\[\comp(\varepsilon,d) \leq  \exp\{p(\phi(\varepsilon^{-1})\psi(d))\}.\]

Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for non-separable tractability
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,e^pd^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)}e^{-p}d^{-p} < \infty \qquad \text{for some} p \geq 0, L_p \in \naturals.
\]
We observe that Example 5 is a special case  where in this case the function $\phi(\varepsilon^{-1})= 1+\log(\max\{1,\varepsilon^{-1}\})$ and $\psi(d) = 1+\log(d)$.
\end{example}
